---
output: github_document
bibliography: inst/REFERENCES.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(weatherQC)
```

# weatherQC

<!-- badges: start -->
<!-- badges: end -->

weatherQC allows standardized quality control checks on historical daily weather
observations including temperature (minimum, maximum, mean) and precipitation (daily
sum).

## Installation

You can install the development version of weatherQC from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("larscaspersen/weatherQC")
```

## How to use the package

weatherQC allows the user to perform standardized quality checks on the 
daily weather data. The goal is to fish out as many not-trustworthy observation
and to keep the invasiveness of the operation as low as possible. The functions
`weather_qc_durre()` and `weather_qc_costa()` are the work horses of the 
package. The functions are built after the described weather quality control schemes by @costa_gap_2021 and @durre_comprehensive_2010. Both contain sets of plausibility checks for the weather data including
basic tests (for example does the observation exceed existing temperature and
rainfall records?), temporal plausibility tests (for example erratic day-to-day changes) to spatial plausibility tests (does the observation make sense in the 
context of neighbouring weather station observations?).

The function call is similar in both cases: the targeted weather needs to be specified, as well as the coordinates of the target weather station. Furthermore the functions
need neighbouring weather station observations for spatial consistency test as
well as the information on the neighbouring stations coordinates. In case of 
`weather_qc_costa` the tested weather variable also needs to be specified. `weather_qc_durre` automatically applies the tests on Tmin, Tmax and Precip columns.

The data of the target weather needs to be supplied via a named list of data.frames
or tibbles. The individual data.frames need to contain the columns
`Year`, `Month`, `Day`, `Tmin`, `Tmax`, `Tmean` and `Precipitation`. An example can
be seen in the data.frame supplied wiht the package:

```{r }
head(target_weather)
```
As you can see, having more columns than the previous mentioned ones is fine.
Furthermore the 
Here are two examples how to call the function.

```{r  eval = F}
## how to use the function
weatherQC::weather_qc_costa(weather_list = weather_list)

weatherQC::weather_qc_durre(weather_list = weather_list)
```

In both cases the function return the same list supplied via the `weather_list` argument with six extra columns, two for each of the tested variables Tmin, Tmax
and Precip. These will be explained at the example of the column for minimum temperature Tmin. The original column of Tmin is altered by the function, if the test lead to a flagging of the observation it is replaced with NA. In case of `weather_qc_durre` one flag is enough for the removal of the observation, which test lead to it is indicated in `flag_Tmin`. `Tmin_org` contains the original, unaltered observations. In case of `weather_QC_durre` at least two tests need to flag an 
observation before it gets replaced by NA. The comments in `flag_Tmin` also 
look a bit different. The listed numbers correspond to the five tests, which are
(1) fixed limit test, (2) variable limit test, (3) temporal consistency test
, (4) consistency among variables (which only works for Tmin and Tmax) and (5) spatial consistency test. In cases of only one listed number, the observation was not removed.

```{r }
#weather data needs to be supplied in a named list
weather_list <- list(target_weather)
names(weather_list) <- target_info$id

#example output
qc_outcome <- weather_qc_costa(weather_list = weather_list, 
                               mute = T)

qc_outcome[[1]][750:760, c("Date", "Tmin", "Tmax", "Tmin_org", "Tmax_org", "flag_Tmin", "flag_Tmax")]

```

The function keeps the user updated on which test it is currently working, but
the function can also be muted with the `mute = TRUE` argument. The spatial
consistency tests involve a lot of computation are currently rather slow. If
the user wishes to skip the spatial consistency tests, then this can be done
with the `skip_spatial_test = TRUE` argument. In case the spatial consistency
tests should be carried out, the `skip_spatial_test` should be set to FALSE and
the arguments `weather_info`, `aux_list` and `aux_list` need to be supplied as well.
`weather_info` and `aux_info` should be both data.frames with as many rows as elements
in `weather_list` and `aux_list`. Furthermore, the columns `id` with a unique
identifier for each weather station and weather station coordinates in decimal 
format should be present in the columns `Longitude` and `Latitude`. An example for
the aux_info can be seen here:

```{r }
head(neighbour_info)

```
The aux_list should be structured like in weather_list, a list of named data.frames (same names as `id` column in aux_info) with the columns Year, Month, Day, Tmin, Tmax and Precip.

```{r }
head(neighbour_weather[[1]])

```

## Help and further ressources
For more details on the individual tests within `weather_qc_costa` and `weather_qc_durre` please use the function help pages which can be found via `?weather_qc_durre` or `help(weather_qc_durre)`. There will be also a further vignette
which explains the inner workings of `weather_qc_costa` and `weather_qc_durre` in more detail.

## References