---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# weatherQC

<!-- badges: start -->
<!-- badges: end -->

weatherQC allows standardized quality control checks on historical daily weather
observations including temperature (minimum, maximum, mean) and precipitation (daily
sum).

## Installation

You can install the development version of weatherQC from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("larscaspersen/weatherQC")
```

## How to use the package

weatherQC allows the user to perform standardized quality checks on the 
daily weather data. The goal is to fish out as many not-trustworthy observation
and to keep the invasiveness of the operation as low as possible. The functions
`weather_qc_durre()` and `weather_qc_costa()` are the work horses of the 
package. Both contain sets of plausibility checks for the weather data including
basic tests (for example does the observation exceed existing temperature and
rainfall records?), temporal plausibility tests (for example erradic day-to-day changes) to spatial plausibility tests (does the observation make sense in the 
context of neighbouring weather station observations?).

```{r  eval = F}

## how to use the function
weatherQC::weather_qc_costa(weather =  target_weather, 
                 weather_coords = c(target_info$Longitude, target_info$Latitude),
                 variable = "Tmin", 
                 aux_list = neighbour_weather,
                 aux_info = neighbour_info)

weatherQC::weather_qc_durre(weather =  target_weather, 
                 weather_coords = c(target_info$Longitude, target_info$Latitude),
                 aux_list = neighbour_weather,
                 aux_info = neighbour_info)
```

Explain the different objects in more detail

```{r }
library(weatherQC)
target_info

```

```{r }
head(neighbour_info)

```

```{r }
head(target_weather)

```

```{r }
head(neighbour_weather[[1]])

```


show output of the function


## details on weather_qc_durre and weather_qc_costa
 show the functions in more details